---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

# ggplot - display colour palette
fun_display <- function(col_pal) {
  n <- length(col_pal)
  
df1 <- data.frame(x = 1:n,
             y = 1,
             col = col_pal)

myColors <- col_pal
names(myColors) <- as.factor(df1$col)

  ggplot(data = df1, aes(x, y , fill = col)) +
    geom_tile() +
    scale_fill_manual(values = myColors) +
    theme_void() +
    theme(legend.position = "none")
}
# 
# library(tidyverse)
# library(gapminder)
 library(CawthronColours)
 library(patchwork)
# library(rnaturalearth)
# 
# nz_map <- ne_states(country = "new zealand", returnclass = "sf")
# 
# nz_map <- nz_map %>%
#   dplyr::filter(region %in% c("South Island", "North Island"))
# 
# nz_map <- nz_map %>%
#   select(name) %>%
#   mutate(count = runif(n(), 0, 100))
```

# CawthronColours <img src="man/figures/cawthronColours_logo.svg" align="right" height="150"/>

<!-- badges: start -->
<!-- badges: end -->

## package in development - could break at any moment ....

The goal of `CawthronColours` is to to provide official Cawthron colour palettes for use in R. The colour palettes are a mix of well known (e.g. `viridis`) and custom palettes. All are colour blind friendly.

Inspiration for package implementation came from the the [`Manu`](https://github.com/G-Thomson/Manu) colour palette package. All palettes except `viridis` and `lawa` were designed by Jacqui Stuart.


## Installation

The package is currently hosted on GitHub and can be installed as follows:

``` {r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("cawthron/CawthronColours")
```

## Usage

The colour palettes are stored as a list named `cawthron_palettes`. There are categorical (`caw_cat_*`), sequential (`caw_seq_*`) and diverging (`caw_div_*`) palettes.You can see a list of the available palettes like so:
```{r example}
library(CawthronColours)
names(cawthron_palettes)
```

A helper function `get_pal()` returns the desired colour palette as a vector, for example:
```{r}
get_pal("caw_cat_1")
```

There is also a helper function `print_pal()` which displays the palette in the graphics window.
```{r, fig.height=1}
caw_col <- get_pal("caw_cat_1")
print_pal(caw_col)
```

Below area few examples of plots made using [`ggplot2`](https://ggplot2.tidyverse.org/). We use the [`gapminder`](https://github.com/jennybc/gapminder) package to get some toy data and [`rnaturalearth`](https://github.com/ropensci/rnaturalearth) to get a map of New Zealand. The key `ggplot2` functions needed to use the `CawthronColours` palettes are:

For categorical palettes:

* `scale_colour_manual()`
* `scale_fill_manual()`


For diverging or sequential palettes:

* `scale_colour_gradientn()`
* `scale_fill_gradientn()`


```{r, message=FALSE, warning=FALSE}
library(gapminder)
library(ggplot2)
library(dplyr)
library(rnaturalearth)

# For categorical colours we use scale_colour_manual (or scale_fill_manual)
ggplot(gapminder, aes(x = continent, y = lifeExp, colour = continent)) +
  geom_jitter(position = position_jitter(width = 0.1, height = 0)) +
  scale_colour_manual(values = get_pal("caw_cat_1")) +
  theme_classic() +
  theme(legend.position = "none")
# for continuous or sequential palettes we use scale_fill_gradientn (or scale_colour_gradientn)
ggplot(faithfuld) +
  geom_tile(aes(waiting, eruptions, fill = density)) +
  scale_fill_gradientn(colours = get_pal("caw_seq_1"))  +
  theme_classic() +
  theme(legend.position = "none")
# download a map of NZ
nz_map <- rnaturalearth::ne_states(country = "new zealand", returnclass = "sf")

# remove outlying islands
nz_map <- nz_map |>
  dplyr::filter(region %in% c("South Island", "North Island"))

# add a random variable for each region
nz_map <- nz_map %>%
  dplyr::select(name) |>
  dplyr::mutate(count = runif(dplyr::n(), 0, 100))

# plot - use scale_fill_gradientn and scale_colour_gradientn to colour the
# regions by the random variable we just created
ggplot(nz_map, aes(fill = count, colour = count)) +
  geom_sf() +
  scale_fill_gradientn(colours = get_pal("caw_seq_1"))  +
  scale_colour_gradientn(colours = get_pal("caw_seq_1"))  +
  theme_void() +
  theme(legend.position = "none")
```

# Palettes
Overview of the different colour palettes 

## Categorical palettes

### Cawthron rainbow - `caw_cat_1`

```{r, echo=FALSE}
p1 <- fun_display(get_pal("caw_cat_1"))

p2 <- ggplot(gapminder, aes(x = continent, y = lifeExp, colour = continent)) +
  geom_jitter(position = position_jitter(width = 0.1, height = 0)) +
  scale_colour_manual(values = get_pal("caw_cat_1")) +
  theme_classic() +
  theme(legend.position = "none")

p1 / p2  + plot_layout(heights = c(1, 3))
```


### Warm and cool - `caw_cat_2`

```{r, echo=FALSE}
p1 <- fun_display(get_pal("caw_cat_2"))

p2 <- ggplot(gapminder, aes(x = continent, y = lifeExp, colour = continent)) +
  geom_jitter(position = position_jitter(width = 0.1, height = 0)) +
  scale_colour_manual(values = get_pal("caw_cat_2")) +
  theme_classic() +
  theme(legend.position = "none")

p1 / p2  + plot_layout(heights = c(1, 3))
```

### Lawa - `caw_cat_3`

```{r, echo=FALSE}
p1 <- fun_display(get_pal("caw_cat_3"))

p2 <- ggplot(gapminder, aes(x = continent, y = lifeExp, colour = continent)) +
  geom_jitter(position = position_jitter(width = 0.1, height = 0)) +
  scale_colour_manual(values = get_pal("caw_cat_3")) +
  theme_classic() +
  theme(legend.position = "none")

p1 / p2  + plot_layout(heights = c(1, 3))
```

## Sequential palettes

### Tuaone - `caw_seq_1`

```{r, echo=FALSE}
p1 <- ggplot(nz_map, aes(fill = count, colour = count)) +
  geom_sf() +
  scale_fill_gradientn(colours = get_pal("caw_seq_1"))  +
  scale_colour_gradientn(colours = get_pal("caw_seq_1"))  +
  theme_void() +
  theme(legend.position = "none")

p2 <- ggplot(faithfuld) +
  geom_tile(aes(waiting, eruptions, fill = density)) +
  scale_fill_gradientn(colours = get_pal("caw_seq_1"))  +
  theme_classic() +
  theme(legend.position = "none")

p0 <- fun_display(get_pal("caw_seq_1"))


p0 / (p1 + p2) + plot_layout(heights = c(1, 3))
```

### Viridis - `caw_seq_2`

```{r, echo=FALSE}
p1 <- ggplot(nz_map, aes(fill = count, colour = count)) +
  geom_sf() +
  scale_fill_gradientn(colours = get_pal("caw_seq_2"))  +
  scale_colour_gradientn(colours = get_pal("caw_seq_2"))  +
  theme_void() +
  theme(legend.position = "none")

p2 <- ggplot(faithfuld) +
  geom_tile(aes(waiting, eruptions, fill = density)) +
  scale_fill_gradientn(colours = get_pal("caw_seq_2"))  +
  theme_classic() +
  theme(legend.position = "none")

p0 <- fun_display(get_pal("caw_seq_2"))


p0 / (p1 + p2) + plot_layout(heights = c(1, 3))
```

### Tāne Mahuta - `caw_seq_3`

```{r, echo=FALSE}
p1 <- ggplot(nz_map, aes(fill = count, colour = count)) +
  geom_sf() +
  scale_fill_gradientn(colours = get_pal("caw_seq_3"))  +
  scale_colour_gradientn(colours = get_pal("caw_seq_3"))  +
  theme_void() +
  theme(legend.position = "none")

p2 <- ggplot(faithfuld) +
  geom_tile(aes(waiting, eruptions, fill = density)) +
  scale_fill_gradientn(colours = get_pal("caw_seq_3"))  +
  theme_classic() +
  theme(legend.position = "none")

p0 <- fun_display(get_pal("caw_seq_3"))


p0 / (p1 + p2) + plot_layout(heights = c(1, 3))
```

### Tangaroa - `caw_seq_4`

```{r, echo=FALSE}
p1 <- ggplot(nz_map, aes(fill = count, colour = count)) +
  geom_sf() +
  scale_fill_gradientn(colours = get_pal("caw_seq_4"))  +
  scale_colour_gradientn(colours = get_pal("caw_seq_4"))  +
  theme_void() +
  theme(legend.position = "none")

p2 <- ggplot(faithfuld) +
  geom_tile(aes(waiting, eruptions, fill = density)) +
  scale_fill_gradientn(colours = get_pal("caw_seq_4"))  +
  theme_classic() +
  theme(legend.position = "none")

p0 <- fun_display(get_pal("caw_seq_4"))

p0 / (p1 + p2) + plot_layout(heights = c(1, 3))
```

### Papatūānuku - `caw_seq_5`

```{r, echo=FALSE}
p1 <- ggplot(nz_map, aes(fill = count, colour = count)) +
  geom_sf() +
  scale_fill_gradientn(colours = get_pal("caw_seq_5"))  +
  scale_colour_gradientn(colours = get_pal("caw_seq_5"))  +
  theme_void() +
  theme(legend.position = "none")

p2 <- ggplot(faithfuld) +
  geom_tile(aes(waiting, eruptions, fill = density)) +
  scale_fill_gradientn(colours = get_pal("caw_seq_5"))  +
  theme_classic() +
  theme(legend.position = "none")

p0 <- fun_display(get_pal("caw_seq_5"))

p0 / (p1 + p2) + plot_layout(heights = c(1, 3))
```


## diverging palettes

### Sun to sea - `caw_div_1`

```{r, echo=FALSE}
p1 <- ggplot(nz_map, aes(fill = count, colour = count)) +
  geom_sf() +
  scale_fill_gradientn(colours = get_pal("caw_div_1"))  +
  scale_colour_gradientn(colours = get_pal("caw_div_1"))  +
  theme_void() +
  theme(legend.position = "none")

p2 <- ggplot(faithfuld) +
  geom_tile(aes(waiting, eruptions, fill = density)) +
  scale_fill_gradientn(colours = get_pal("caw_div_1"))  +
  theme_classic() +
  theme(legend.position = "none")

p0 <- fun_display(get_pal("caw_div_1"))

p0 / (p1 + p2) + plot_layout(heights = c(1, 3))
```


### Hot to cold - `caw_div_2`

```{r, echo=FALSE}
p1 <- ggplot(nz_map, aes(fill = count, colour = count)) +
  geom_sf() +
  scale_fill_gradientn(colours = get_pal("caw_div_2"))  +
  scale_colour_gradientn(colours = get_pal("caw_div_2"))  +
  theme_void() +
  theme(legend.position = "none")

p2 <- ggplot(faithfuld) +
  geom_tile(aes(waiting, eruptions, fill = density)) +
  scale_fill_gradientn(colours = get_pal("caw_div_2"))  +
  theme_classic() +
  theme(legend.position = "none")

p0 <- fun_display(get_pal("caw_div_2"))

p0 / (p1 + p2) + plot_layout(heights = c(1, 3))
```



### Whenua ki te wai - `caw_div_3`

```{r, echo=FALSE}
p1 <- ggplot(nz_map, aes(fill = count, colour = count)) +
  geom_sf() +
  scale_fill_gradientn(colours = get_pal("caw_div_3"))  +
  scale_colour_gradientn(colours = get_pal("caw_div_3"))  +
  theme_void() +
  theme(legend.position = "none")

p2 <- ggplot(faithfuld) +
  geom_tile(aes(waiting, eruptions, fill = density)) +
  scale_fill_gradientn(colours = get_pal("caw_div_3"))  +
  theme_classic() +
  theme(legend.position = "none")

p0 <- fun_display(get_pal("caw_div_3"))

p0 / (p1 + p2) + plot_layout(heights = c(1, 3))
```


