---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CawthronColours

<!-- badges: start -->
<!-- badges: end -->

## package in development - not yet functional!

The goal of CawthronColours is to to provide official Cawthron colour palettes for use in R. The colour palettes are a mix of well known (e.g. viridis) and custom palettes. All are colour blind friendly.

## Installation

You can install the development version of CawthronColours like so:

``` {r, eval = FALSE}
if (!require("CawthronColours")) install.packages("CawthronColours")
remotes::install_gitlab("cinst/cawthron-internal-tools/CawthronColours")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(CawthronColours)


```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
